---
title: "site1_accuracy"
author: "Natalie Woo"
date: "2024-12-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, message=F}
library(mlr3)
library(mlr3learners)
library(mlr3spatial)
library(terra)
library(sf)
library(ggplot2)
library(tidyterra)
library(here)
```

# Import Data
Import accuracy polygons and classified raster 
```{r}
site1_poly <- vect(here("Data", "site1", "site1_accuracy_poly.shp"))
site1_classified <- rast(here("Data", "site1", "site1_classified.tif"))
```

# Create points at the same locations containing information of the correct class and classified class
"correct" class is provided by the "accuracy_pts" dataset, which contains points from polygons we drew ourselves
```{r}
# Create empty spatvector to store points
site1_accuracy_pts <- vect()
```

```{r}
#Loop through each polygon in the shapefile and create 200 random points in each one 
set.seed(103)
for (i in 1:nrow(site1_poly)) {
  single_poly <- site1_poly[i, ] #look at each individual polygon within "site1_poly" 
  randomPts <- spatSample(single_poly, 200, method = "random") #create 200 random points 
  site1_accuracy_pts <- rbind(site1_accuracy_pts, randomPts) #store the random points in the empty spatVector created previously
}
print(site1_accuracy_pts)
```

```{r}
#Save points as a shapefile
writeVector(site1_accuracy_pts, here("Data", "site1", "site1_accuracy_pts.shp"), filetype = "ESRI Shapefile", overwrite=T) #can omit
```

"classified_pts" contains information on which class the points were classified as using a random forest classifier
```{r}
#Extract classes from the classified raster, at the random points previously generated
site1_classified_pts <- extract(site1_classified, site1_accuracy_pts)
```

```{r}
#obtain classes from the classified raster and premade points 
site1_classified_pts <- site1_classified_pts[,2]
site1_accuracy_pts <- site1_accuracy_pts$class
```

```{r}
#Calculate the classification accuracy
##find the number of classified points that match the correct points
site1_accuracy_vect <- site1_classified_pts == site1_accuracy_pts 
##divide the number of points which match by the total number of points
site1_accuracy <- sum(site1_accuracy_vect) / length(site1_accuracy_vect) 
print(paste("Classification Accuracy for site1:", site1_accuracy))
```



